---
title: "Assignment1"
author: "Roland Pfeifer"
date: "18 November 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

Dataset: [activity.zip](<https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip>).

### 0. Load required libraries
```{r}
library(ggplot2)
library(dplyr)
library(chron)
```


### 1. Code for reading in the dataset and processing the data
```{r, eval=TRUE}
activity <- read.csv("activity.csv", header = TRUE)
```

### 2. Histogram of the total number of steps taken each day
```{r, eval=TRUE}
numberofsteps<- aggregate(steps ~ date, activity, FUN=sum)
hist(numberofsteps$steps, 
     col="green", 
     xlab = "Frequency", 
     ylab = "Steps",
     main = "Total Number Of Steps Taken Each day")
```

### 3. Mean and median number of steps taken each day
```{r, eval=TRUE}
amean <- mean(numberofsteps$steps)
amedian <- median(numberofsteps$steps)
amean
amedian
```

Mean and Median total number of steps taken per day are `r format(amean, scientific=FALSE)` and `r format(amedian, scientific=FALSE)` respectively.


### 4. Time series plot of the average number of steps taken
```{r, eval=TRUE}
agginterval <- aggregate(steps ~ interval, activity, FUN=mean)
plot(agginterval$interval, agginterval$steps, 
     type = "l", lwd = 2,
     xlab = "Interval", 
     ylab = "Avg Steps",
     main = "Average Steps vs. 5-Minute Interval")
```

### 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r, echo=TRUE}
filter(agginterval, steps==max(steps))
```


### 6. Code to describe and show a strategy for imputing missing data
```{r}
table(is.na(activity))
meaninterval<- aggregate(steps ~ interval, activity, FUN=mean)
anew <- merge(x=activity, y=meaninterval, by="interval")
anew$steps <- ifelse(is.na(anew$steps.x), anew$steps.y, anew$steps.x)
head(anew)
anew <- select(anew, steps, date, interval)
head(anew)

```

### 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
aggsteps_new<- aggregate(steps ~ date, anew, FUN=sum)
par(mfrow=c(1,2))
hist(aggsteps_new$steps, 
     col="green",
     xlab = "Steps", 
     ylab = "Frequency",
     ylim = c(0,35),
     main = "Total Steps per day",
     cex.main = 0.7)
```

### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
table(is.weekend(anew$date))
anew$dayofweek <- ifelse(is.weekend(anew$date), "weekend", "weekday")
table(anew$dayofweek)
head(anew)
meaninterval_new<- aggregate(steps ~ interval + dayofweek, anew, FUN=mean)
head(meaninterval_new)
ggplot(meaninterval_new, aes(x=interval, y=steps)) + 
  geom_line(color="blue", size=1) + 
  facet_wrap(~dayofweek, nrow=2) +
  labs(x="\nInterval", y="\nNumber of steps")
```


